---
title: "Behavioral output paper 1"
author: "Pieter Huycke"
date: "19/11/2020"
output: pdf_document
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = F)

setwd("C:/Users/pieter/OneDrive - UGent/Projects/2019/overtraining - PILOT 3/figures/Publish/Data/Stimulus-locked/Theta, alpha, beta + behavioral data")
df = read.csv("theta_alpha_beta_behavioural.csv", 
              header = TRUE, sep = ",")
```

## General data manipulation

In the following block, we do some general manipulation on the dataset. We first check the data to see how it looks. Then we load in the relevant libraries for analysis. Finally, we convert some variables to factors.

```{r, label='general-manipulations', echo=F, results='hide',message=F}
# check if data load worked out
head(df)

# load needed libraries
library(car)
library(lme4)

# VARIABLE MANIPULATION  ---------

# see as certain predictors as factor
df$Block_overall       = factor(df$Block_overall)
df$Block_specific      = factor(df$Block_specific)
df$Repetitions_overall = factor(df$Repetitions_overall)
df$Repetitions_block   = factor(df$Repetitions_block)
df$Subject_nr          = factor(df$Subject_nr)

# set the Anova parameters
options(contrasts = c("contr.sum", "contr.poly"))
```

## Slow timescale

In the next part, we look at the behavioral effects on the slow timescale. Note that slow timescale means that we look at effects over experimental blocks. Thus, our independent variables are block number (accounting for condition), and condition. Block numbers within a condition range from 1 to 8 (both included), condition is binary ('novel', or 'recurring'). A condition that is 'novel' means that the stimuli are never seen before, and are unique to this block. 'Recurring' blocks show stimuli (and their associated mappings) that repeat throughout the experiment. We look at variations in both reaction times (RT) and errors made. For both dependent variables, we investigate the main effect of block number and condition, and their interaction effect. We use the functiona `lmer()` and `glmer` to accomplish this, respectively.

```{r, label='aov1', echo=T}
# log(RT) ~ block number + condition + (block number x condition) 
rt.block_cond = lmer(RT_log  ~ (1|Subject_nr) + Block_specific * Condition, 
                     data = df)
aov1 = Anova(rt.block_cond,
             type           = "III",
             test.statistic = "F")
aov1
```

```{r, label='aov2', echo=T}
# error rate ~ block number + condition + (block number x condition) 
acc.block_cond = glmer(Error_int ~ (1|Subject_nr) + Block_specific * Condition, 
                       data    = df,
                       family  = binomial,
                       control = glmerControl(optimizer = "Nelder_Mead"))
aov2 = Anova(acc.block_cond,
             type           = "III",
             test.statistic = "Chisq")
aov2
```

## Fast timescale

In the subsequent analysis, we look at the effects of stimulus repetition on both RT and error rate. Note that we first do an extra manipulation: we subset the data to make sure that the maximum amount of stimulus repetitions is 8. This makes sure that we can compare between conditions, since both the novel- and the recurring condition will have a maximal repetition occurence of 8. We only look at the main effect of stimulus repetition, not at the main effect of condition.

```{r, label='aov3', echo=T}
# make a subset for when repetitions <= 8
df.reduced = df[as.numeric(as.character(df$Repetitions_overall)) <= 8, ]

# log(RT) ~ repetitions 
rt.reps = lmer(RT_log  ~ (1|Subject_nr) + Repetitions_block, 
               data = df.reduced)
aov3 = Anova(rt.reps,
             type           = "III",
             test.statistic = "F")
aov3
```

``````{r, label='aov4', echo=T}
# error rate ~ repetitions 
acc.reps = glmer(Error_int ~ (1|Subject_nr) + Repetitions_block, 
                 data    = df.reduced,
                 family  = binomial,
                 control = glmerControl(optimizer = "Nelder_Mead"))
aov4 = Anova(acc.reps ,
             type           = "III",
             test.statistic = "Chisq")
aov4
```

